### 2015-07-05 narrative/qos.tex
modified:   qos.tex

\paragraph{Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
Performance quality can be expressed in either relative or absolute
terms. Examples of relative performance terms are ``fair'',
``proportional'', or ``priority/class-based'' (``the higher the
priority the better the service'' or ``1st class is better than
2nd''). The key advantage of relative performance terms is that
they are easy to implement. The key disadvantage is the difficulty
of making any guarantees based on those implementations --
even relative guarantees are mired by the well-known effect of
priority inversion~\cite{lampson:cacm80}. Examples of absolute
performance terms are ``rate-based'', ``soft real-time'', and
``hard real-time''. The key advantage of absolute performance
terms are that they enable the implementation of strong
guarantees: absolute terms enable agreements between application
tasks and storage that we call \emph{reservations} and that do not
change meaning depending
on the storage system's workload. The key challenge of absolute
performance terms is that their implementation is non-trivial,
especially in large-scale systems.

>>>>

\paragraph{Resource Management Background:} There is an important connection between finding the resources that can help fulfill a particular request and the predictability of the performance of that request, both in terms of latency and resolution. This connection is particular important for reads: in very large scale systems, finding a resource to read can be a significant, if not dominant, component of overall read latency. If the search for resources is unbounded, it will be difficult to make any performance predictions.

Locating resources is a typical distributed systems problem. Central solutions are limited in their scalability, especially if they have to track every update. Traditional file systems are good examples for this: most updates to files require updates to their centrally maintained block allocation tables. The opposite of file systems are ``pile'' systems which instead of keeping track of updates search for resources at request time. Pile systems allow very efficient writes because clients can write whereever there is free space and there is very bookkeeping overhead. Furthermore, pile systems have great flexibility for moving resources around, for example to ensure durability. However, reads require a potentially unbounded search, especially if it is not clear whether the resource actually exists. Methods to make searches more efficient without reverting to centralized bookkeeping usually involves some kind of partitioning of the search space and a method to limit the search to a small number of partitions.

Frequently this involves a hash function that is relatively stable against changes to its range, such as failures or updates to the hardware of the system (see for example consistent hashing~\cite{karger:stoc97}). Consistent hashing is scalable since it can be well-known throughout a system of any size and only requires minimal movement of data in the presence of failures or system changes. Consistent hash functions can also have sets of ordered but infinite lists as their range (which works well with primary replication)~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support placement policy rules~\cite{weil:sc06}.

In this proposal we will investigate the trade-off in terms of predictable peformance between ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' data, pushing the cost of finding to the moment when the data is actually needed. This trade-off has a parallel in data management where there are two alternative paths to achieve knowledge: the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront cost and query the raw data directly by using data processing platforms such as Hadoop. Recent work has found that the combination of these two techniques can be optimized to perform better than any one technique alone~\cite{lefevre:sigmod14a}. We believe that a similar opportunity exists for combining filing and piling approaches to achieve performance predictability without sacrificing too much flexibility.

More specifically, the coarseness of a search space partitioning limits the search time within a partition, i.e. finely partitioned search spaces require little or no search within a partition but offer little placement flexibility. Partitioning can also be fine in one dimension and coarse in another to allow flexibility where it is most needed while still ensuring fast finding. For example, a hash function can finely partition a low storage tier but each partition can extend across multiple tiers so that a search still has to occur in order to find a resource on a particular tier.

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
Performance quality can be expressed in either relative or absolute
terms. Examples of relative performance terms are ``fair'',
``proportional'', or ``priority/class-based'' (``the higher the
priority the better the service'' or ``1st class is better than
2nd''). The key advantage of relative performance terms is that
they are easy to implement. The key disadvantage is the difficulty
of making any guarantees based on those implementations --
even relative guarantees are mired by the well-known effect of
priority inversion~\cite{lampson:cacm80}. Examples of absolute
performance terms are ``rate-based'', ``soft real-time'', and
``hard real-time''. The key advantage of absolute performance
terms are that they enable the implementation of strong
guarantees: absolute terms enable agreements between application
tasks and storage that we call \emph{reservations} and that do not
change meaning depending
on the storage system's workload. The key challenge of absolute
performance terms is that their implementation is non-trivial,
especially in large-scale systems.
